Multi-Resolution Neural Field Estimation

### 2011-03-12 Multiresolution_framework/ltx/manuscript.tex
edit to research highlights

The actual connectivity kernel and the decomposition is plotted in \figurename{\ref{fig:KernelEstimate}}(a). No assumptions where made about the shape of the kernel. The reconstructed kernel is in good accordance with the actual kernel, where the actual kernel lies inside the confidence interval. The large standard deviation is due to the high number of parameters need to be estimated. The small error in the estimate is attributed to the MRA of the system used to form the estimator. \figurename{\ref{fig:KernelEstimate}}(b) and (c) shows the kernel scaling and wavelet functions, respectively.
\begin{figure}[!h] 
	\centering
		\includegraphics[scale=1]{./Graph/fig1.pdf}
		\caption{Actual (gray) and estimated (black) neural fields at a time instant for different spatial resolutions. (a) $j=0$, $n_x=17$, $RMSE = 1.93$. (b) $j=1$, $n_x=33$, $RMSE = 1.39$. (c) $j=2$, $n_x=65$, $RMSE = 0.90$. (d) $j=3$, $n_x=131$, $RMSE = 0.78$. (e) $j=4$, $n_x=263$, $RMSE=0.78$.}
	\label{fig:FieldEstimates}
\end{figure} 
\begin{figure}[!h] 
	\centering
		\includegraphics[scale=1]{./Graph/fig2.pdf}
		\caption{\parham{Connectivity kernel estimate. (a) The actual kernel and its components are shown by solid and dashed black lines respectively. The mean kernel estimates over 200 realizations and confidence interval are shown by the red line and red shaded region ($\pm2$ std.). (b)-(c) Kernel scaling and wavelets functions, respectively, at $j=1$ used in estimation algorithm.}}
	\label{fig:KernelEstimate}
\end{figure}
% \caption{The actual connectivity kernel is shown with the black solid line. The estimated kernel and confidence interval is shown by the red line and red shaded region ($\pm2$ std.). The 7 weighted kernel basis functions are shown by the black dashed lines.}
\section{Discussion}
In this paper, we have presented a novel model-based framework for estimating cortical dynamics from electrophysiological measurements. The novel and key developments of the paper include the multi-resolution FEM representation of the neural field, and the estimator for an intracortical connectivity kernel with an arbitrary shape. This work is significant, as the ability to create patient specific neural field models has the potential to contribute to our understanding and improve treatment of diseases resulting from abnormal neurodynamics, such as epilepsy. Other groups have previously highlighted the importance of a multi-resolution approach in neural field modeling. It is thought that the dynamics and connectivity structure differs at different spatial scales~\cite{Qubbaj2009,Breakspear2005}. %\dean{This is a test to see if we can fit something about Breaky's paper in here. It would be nice to slot something in here because he used wavelets.}

@@

In order to apply the framework to real data some assumptions must be made. A critical assumption is that the model provides an apt description of the cortical dynamics. The authors acknowledge that there is, and will always be, a discrepancy between the model and cortex. Nevertheless, the model-based framework proposed in this paper may enable meaningful state tracking and connectivity estimation. The key development is the multi-resolution decomposition forming the state-space model. While the decomposition holds for more sophisticated models, efficiently performing nonlinear smoothing in the large state-space remains a challenge. Additional avenues for future work include extending the model to capture a $2^{nd}$-order synaptic response kernel, and time delays according to spatial resolutions. The estimation techniques should also be extended to deal with spatiotemporal heterogeneity in the kernel. Finally, future work should be directed towards applying and validating the framework on real data.

>>>>

The actual connectivity kernel and the decomposition is plotted in \figurename{\ref{fig:KernelEstimate}}(a). No assumptions where made about the shape of the kernel. The reconstructed kernel is in good accordance with the actual kernel, where the actual kernel lies inside the confidence interval. The large standard deviation is due to the high number of parameters need to be estimated. The small error in the estimate is attributed to the MRA of the system used to form the estimator. \figurename{\ref{fig:KernelEstimate}}(b) and (c) shows the kernel scaling and wavelet functions, respectively.
\begin{figure}[!h] 
	\centering
		\includegraphics[scale=1]{./Graph/fig1.pdf}
		\caption{Actual (gray) and estimated (black) neural fields at a time instant for different spatial resolutions. (a) $j=0$, $n_x=17$, $RMSE = 1.93$. (b) $j=1$, $n_x=33$, $RMSE = 1.39$. (c) $j=2$, $n_x=65$, $RMSE = 0.90$. (d) $j=3$, $n_x=131$, $RMSE = 0.78$. (e) $j=4$, $n_x=263$, $RMSE=0.78$.}
	\label{fig:FieldEstimates}
\end{figure} 
\begin{figure}[!h] 
	\centering
		\includegraphics[scale=1]{./Graph/fig2.pdf}
		\caption{\parham{Connectivity kernel estimate. (a) The actual kernel and its components are shown by solid and dashed black lines, respectively. The mean kernel estimates over 200 realizations and confidence interval are shown by the red line and red shaded region ($\pm2$ std.). (b)-(c) Kernel scaling and wavelets functions, respectively, at $j=1$ used in estimation algorithm.}}
	\label{fig:KernelEstimate}
\end{figure}
% \caption{The actual connectivity kernel is shown with the black solid line. The estimated kernel and confidence interval is shown by the red line and red shaded region ($\pm2$ std.). The 7 weighted kernel basis functions are shown by the black dashed lines.}
\section{Discussion}
In this paper, we have presented a novel model-based framework for estimating cortical dynamics from electrophysiological measurements. The novel and key developments of the paper include the multi-resolution FEM representation of the neural field, and the estimator for an intracortical connectivity kernel with an arbitrary shape. This work is significant, as the ability to create patient specific neural field models has the potential to contribute to our understanding and improve treatment of diseases resulting from abnormal neurodynamics, such as epilepsy. Other groups have previously highlighted the importance of a multi-resolution approach in neural field modeling. It is thought that the dynamics and connectivity structure differs at different spatial scales~\cite{Qubbaj2009,Breakspear2005}. %\dean{This is a test to see if we can fit something about Breaky's paper in here. It would be nice to slot something in here because he used wavelets.}

@@

% In order to apply the framework to real data some assumptions must be made. A critical assumption is that the model provides an apt description of the cortical dynamics. 
The authors acknowledge that there is, and will always be, a discrepancy between the model and cortex. Nevertheless, the model-based framework proposed in this paper may enable meaningful state tracking and connectivity estimation. The key development is the multi-resolution decomposition forming the state-space model. While the decomposition holds for more sophisticated models, efficiently performing nonlinear smoothing in the large state-space remains a challenge. Additional avenues for future work include extending the model to capture a $2^{nd}$-order synaptic response kernel, and time delays according to spatial resolutions. The estimation techniques should also be extended to deal with spatiotemporal heterogeneity in the kernel. Finally, future work should be directed towards applying and validating the framework on real data.
